---
title: "NBA Game Analysis"
author: "Samuel Ivanecky"
date: "2/14/2020"
output: html_document
---

# Setup
```{r}
library(data.table)
library(tidyverse)
library(tidymodels)
```

# Load Data
```{r}
nba = read.csv("nba.games.stats.csv")
```

# Data Cleaning
## Print structure of the data
```{r}
str(nba)
```

## Count number of missing in data
```{r}
# No missing data
sum(is.na(nba))
```

## Rename columns
```{r}
names(nba) = c("index", "team", "game", "date", "home", "opponent", "win_loss", "team_pts", "opp_pts", "fg_made", "fg_atmpt", "fg_pct", "three_pt_made", 
               "three_pt_atmpt", "three_pt_pct", "ft_made", "ft_atmpt", "ft_pct", "off_reb", "tot_reb", "assists", "steals", "blocks", "turnovers", "total_fouls", 
               "opp_fg_made", "opp_fg_atmpt", "opp_fg_pct", "opp_three_pt_made", "opp_three_pt_atmpt", "opp_three_pt_pct", "opp_ft_made", "opp_ft_atmpt", 
               "opp_ft_pct", "opp_off_reb", "opp_tot_reb", "opp_assists", "opp_steals", "opp_blocks", "opp_turnovers", "opp_tot_fouls")

# Reprint for validation
str(nba)
```

# Feature Engineering
## Season
Need to create a season variable that will group games into seasons based on the date of the game.
```{r}
# note that each season "ends" in April
# Convert date to a date object
nba$date = lubridate::ymd(nba$date)

# Create seasons var
nba = nba %>%
  mutate(
    season = case_when(
      date > '2014-01-01' & date < '2015-05-01' ~ '14-15',
      date > '2015-01-01' & date < '2016-05-01' ~ '15-16',
      date > '2016-01-01' & date < '2017-05-01' ~ '16-17',
      date > '2017-01-01' & date < '2018-05-01' ~ '17-18',
      T ~ 'OTHER'
    )
  )

# Convert to factor
nba$season = as.factor(nba$season)

# Print levels
levels(nba$season)
```

## Distance Traveled
Would be good to have a distance traveled var that calculates how far a team had to go for an away game

## Win Pct
Would be good to have team's winning pct at that point in time
```{r}
# Vectorize all the components
```


## Opp Win Pct
Similarly, would be useful to have opponents winning pct




