---
title: "NBA Game Predictions"
output: html_document
---
# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries
library(tidymodels)
library(dplyr)
library(forecast)
```

# Load Data
```{r}
# Read in data
nba <- read.csv("nba.games.stats.csv")
```

# Clean & Organize Data
```{r}
# Print structure of the data
str(nba)

# Convert Date to date and WINorLOSS to binary
nba <- nba %>% 
  mutate(
    Date = lubridate::ymd(Date),
    WINorLOSS = case_when(
      WINorLOSS == 'W' ~ 1,
      T ~ 0
    )
    )
```

## Separate Data by Season
```{r}
# Break into four seasons
nba15 <- nba %>% filter(Date < '2015-07-01')
nba16 <- nba %>% filter(Date > '2015-07-01' & Date < '2016-07-01')
nba17 <- nba %>% filter(Date > '2016-07-01' & Date < '2017-07-01')
nba18 <- nba %>% filter(Date > '2017-07-01')
```

## Rolling Functions
```{r}
# Function to subset each season
getRollingTeams <- function(df){
  # Select only required columns
  df <- df %>% select(Team, Game, Date, WINorLOSS, TeamPoints, OpponentPoints, FieldGoals, FieldGoalsAttempted, FieldGoals., 
                      X3PointShots, X3PointShotsAttempted, X3PointShots., FreeThrows, FreeThrowsAttempted, FreeThrows., 
                      OffRebounds, TotalRebounds, Assists, Steals, Blocks, Turnovers, TotalFouls)
  
  # Create new variables: DefRebounds, PtDifferential
  df <- df %>%
    mutate(
      DefRebounds = TotalRebounds - OffRebounds,
      PtDifferential = TeamPoints - OpponentPoints
    )
  
  # Create rolling variables
  teams <- df %>%
    group_by(Team) %>%
    mutate(
      Wins = cumsum(WINorLOSS),
      #GamesPlayed = count(Game),
      TeamPoints = cummean(TeamPoints),
      PtDifferential = cummean(PtDifferential),
      FieldGoals = cummean(FieldGoals),
      FieldGoalsAttempted = cummean(FieldGoalsAttempted),
      FieldGoals. = cummean(FieldGoals.),
      X3PointShots = cummean(X3PointShots),
      X3PointShotsAttempted = cummean(X3PointShotsAttempted),
      X3PointShots. = cummean(X3PointShots.),
      FreeThrows = cummean(FreeThrows),
      FreeThrowsAttempted = cummean(FreeThrowsAttempted),
      FreeThrows. = cummean(FreeThrows.),
      OffRebounds = cummean(OffRebounds),
      DefRebounds = cummean(DefRebounds),
      TotalRebounds = cummean(TotalRebounds),
      Assists = cummean(Assists),
      Steals = cummean(Steals),
      Blocks = cummean(Blocks),
      Turnovers = cummean(Turnovers),
      TotalFouls = cummean(TotalFouls)
    ) %>%
    mutate(
      Losses = Game - Wins,
      WinPct = Wins / Game,
      JoinDate = lead(Date) # Variable to join to later data. Averages are technically for AFTER the game happened
    )
  
  return(teams)
}
```

## Get Rolling Team Averages
```{r}
rolling15 <- getRollingTeams(nba15)
rolling16 <- getRollingTeams(nba16)
rolling17 <- getRollingTeams(nba17)
rolling18 <- getRollingTeams(nba18)
```

## Subset Overall Games
```{r}
games <- nba %>%
  select(Team, Game, Date, Opponent, WINorLOSS)
```

## Create Team & Opponent Join Tables
```{r}
# Need one table to join TEAM stats and one to join OPPONENT stats
# DF to hold all rolling stats
rollingStats <- rbind(rolling15, rolling16, rolling17, rolling18)

# Create two dataframes
teams <- rollingStats
colnames(teams) <- paste("TEAM", colnames(teams), sep = "_")
opps <- rollingStats
colnames(opps) <- paste("OPP", colnames(opps), sep = "_")

```

## Join TEAM & OPP Stats
```{r}
games <- games %>%
  left_join(teams, by = c("Team" = "TEAM_Team", "Date" = "TEAM_JoinDate")) %>%
  left_join(opps, by = c("Opponent" = "OPP_Team", "Date" = "OPP_JoinDate"))
```

## Drop Any NA Rows
```{r}
games <- na.omit(games)
```

## Drop Unused Columns
```{r}
drop.cols <- c("TEAM_WINorLOSS", "OPP_WINorLOSS", "TEAM_Date", "OPP_Date", "TEAM_Game", "OPP_Game", "Team", "Game", "Date", "Opponent")

games <- games %>%
  select(-one_of(drop.cols))
```

## Convert WINorLOSS back to W/L
```{r}
games <- games %>%
  mutate(
    WINorLOSS = case_when(
      WINorLOSS == 1 ~ 'W',
      T ~ 'L'
    )
  ) %>%
  mutate(
    WINorLOSS = as.factor(WINorLOSS)
  )
```


# Exploratory Analysis
## Winners - Home vs Away
```{r}
# Plot breakout of winners
ggplot(games, aes(WINorLOSS, fill = WINorLOSS)) +
  geom_bar(stat = "count") +
  ggtitle("Distribution of Winners") +
  labs(x = "Winner", y = "Count")
```

# Modeling 
## Split Data into Train/Test
```{r}
# Shuffle data
games <- games[sample(nrow(games), nrow(games)), ]

# Create a split index
split_ind <- floor(0.75*nrow(games))

# Split into training and testing
train <- games[1:split_ind, ]
test <- games[(split_ind+1):nrow(games), ]
```

### Plot WINNER for both train & test
```{r}
ggplot(train, aes(WINorLOSS, fill=WINorLOSS)) +
  geom_bar(stat="count") +
  ggtitle("Distribution of Training Winners")

ggplot(test, aes(WINorLOSS, fill=WINorLOSS)) +
  geom_bar(stat="count") +
  ggtitle("Distribution of Testing Winners")
```

## Basic Tree Model
### Load Libraries
```{r}
library(rpart)
library(party)
library(maptree)
library(partykit)
```

### Create Model
```{r}
basicTree <- rpart(WINorLOSS ~ ., data = train, method = 'class')

# Plot tree
draw.tree (basicTree, cex=1.1, 
           nodeinfo=TRUE,
           cases="obs",
           digits=1, print.levels=TRUE,
           new=TRUE)
```

### Generate Predictions using Basic Tree
```{r}
# Generate preds
preds <- predict(basicTree, test, type = 'class')

# Compare to actual values
actuals <- test$WINorLOSS

correct <- 0

for (i in 1:length(actuals))
{
  if(actuals[i] == preds[i])
  {
    correct = correct + 1
  }
}

# Print prediction accuracy
print(correct / length(actuals))
```

### Random Forest
```{r}
library(randomForest)

rf <- randomForest(WINorLOSS ~ ., data=train)

pred <- predict(rf, newdata = test[, c(2:47)])

cm <- table(test[,1], pred)
```

# Tidymodels
## Split Data Into Train/Test
```{r}
# Split data on 70/30 ratio
games_split <- initial_split(games, prop = 0.7)
```
